### 2017-06-07 acknowledgments.tex

\section{Acknowledgments}

UPDATE THIS acknowledgments

Javier Vargas is the recipient of a ‘‘Comfuturo” grant funded by
the Fundación General CSIC, C.O.S. Sorzano is the recipient of a
Ramón y Cajal fellowship from the Spanish Ministry of Economy
and Competitiveness (MINECO). The authors acknowledge financial
support from the MINECO through Grants AIC-A-2011-0638,
BFU2013-41249-P and BIO2013-44647-R, the Comunidad de
Madrid through Grant CAM (S2010/BMD-2305), the European
Union (EU) and Horizon 2020 through Grants EINFRA-2014-2:
EGI Engage (Proposal: 654142) and EINFRA-2015-1: West-Life
(Proposal: 675858).
This work was partially funded by Instruct, a component of the
European Strategy Forum on Research Infrastructures (ESFRI), and
supported by national member subscriptions.

    AFTER THIS LINE I JUST SAVED SOME TEXT THAT MAY BE USEFUL IN THE FUTURE DISREGARD THIS MATERIAL
   --------------------

\section{Getting Started}

\section{Project stability and funding. getting started}

Scipion and ScipionBox development is been done in the context of the Instruct Image Processing Cen-
ter (I2PC, http://i2pc.cnb.csic.es). Instruct is the European initiative for Strategic
Scientific Infrastructures (ESFRI) for Structural Biology. The I2PC started with the objective 
to provide efficient support for research projects in Europe demanding expertise in 
electron microscopy image processing. In order to support a large collection
of projects, the first activities of the I2PC were oriented to design and implement the
necessary software tools.

At present the I2PC has this money and this people

ScipionBox is part of the Scipion package which is provided 
freely as open source software. Online documentation
describing Scipion download and installation is available
at \url{http://scipion.cnb.csic.es/m/download_form/} and \url{https://github.com/I2PC/scipion/wiki/How-to-Install}, respectively.

%It was recognized the need for a platform that could provide
%a friendly GUI on top of advanced image processing algorithms and, at the same time,
%assure standardization, traceability and reproducibility.
\section{Things to mention if possible}
\begin{itemize}
 \item Cloud
\end{itemize}