### 2017-06-07 scipionbox.tex

\documentclass[12pt]{article}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{natbib}
\usepackage[T1]{fontenc}
\usepackage[table]{xcolor} % Colors and tables
\usepackage{color, colortbl} % Define new colors and use in tables
\usepackage{setspace} % Set paragraph lines space
%\usepackage{tabu} % Other table package
\usepackage{url} % Other table package
\usepackage{listings}
\usepackage{rotating}
\usepackage[strict]{changepage}
\usepackage[caption=false]{subfig}
\usepackage{fancyhdr}
\usepackage{rotating}
\usepackage{xr}
\usepackage{xspace}
%\externaldocument[Supp-]{supplementary_material}

%\usepackage{lineno}
\usepackage[update,prepend]{epstopdf}
%\linenumbers
\definecolor{lightgreen}{rgb}{0.75,0.75,0.0}
%\usepackage{caption}
%\usepackage{endfloat}
%\DeclareCaptionFormat{empty}{}
%\captionsetup{format=empty}
\pagestyle{fancy}
\fancyhf{} % clear all header and footer fields
\fancyfoot[R]{\footnotesize  \thepage}
\renewcommand{\headrulewidth}{0pt}
\renewcommand{\footrulewidth}{0pt}
\def\scipionbox{\textit{ScipionBox}\xspace}
\def\scipion{\textit{Scipion}\xspace}
\newcommand{\protocol}[1]{\texttt{#1}}

%%%%%%%%%%%%%%%%%\def\baselinestretch{2}
%\def\PARstart#1#2{#1#2} % if draft, disable it
\def\degree{\ensuremath{^\circ}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Textclass specific LaTeX commands.
\newcommand{\lyxaddress}[1]{
   \par {\raggedright #1
   \vspace{1.4em}
   \noindent\par}
}

\newcommand{\itemph}[1]{\item{\emph{#1}}}
\newcommand{\itprog}[1]{\item{\texttt{#1}}:}
\newcommand{\refig}[1]{Fig. \ref{#1}}
\newcommand{\seefig}[1]{(see \refig{#1})}
\definecolor{grey}{rgb}{0.9,0.9,0.9}
\renewcommand{\arraystretch}{1.3}

\lstset{ %
  language=XML,                % the language of the code
  basicstyle=\footnotesize{8pt},       % the size of the fonts that are used for the code
  %numbers=left,                   % where to put the line-numbers
  %numberstyle=\tiny\color{gray},  % the style that is used for the line-numbers
  %stepnumber=2,                   % the step between two line-numbers. If it's 1, each line 
                                  % will be numbered
  %numbersep=5pt,                  % how far the line-numbers are from the code
  backgroundcolor=\color{grey},      % choose the background color. You must add \usepackage{color}
  %showspaces=false,               % show spaces adding particular underscores
  %showstringspaces=false,         % underline spaces within strings
  %showtabs=false,                 % show tabs within strings adding particular underscores
  frame=single,                   % adds a frame around the code
  %rulecolor=\color{black},        % if not set, the frame-color may be changed on line-breaks within not-black text (e.g. commens (green here))
  %tabsize=2,                      % sets default tabsize to 2 spaces
  %captionpos=b,                   % sets the caption-position to bottom
  %breaklines=true,                % sets automatic line breaking
  %breakatwhitespace=false,        % sets if automatic breaks should only happen at whitespace
  %title=\lstname,                   % show the filename of files included with \lstinputlisting;
                                  % also try caption instead of title
  keywordstyle=\color{blue},          % keyword style
  %commentstyle=\color{dkgreen},       % comment style
  %stringstyle=\color{mauve},         % string literal style
  %escapeinside={\%*}{*)},            % if you want to add a comment within your code
  %morekeywords={*,...}               % if you want to add more keywords to the set
}

\begin{document}
\title{ScipionBox: automatic processing for }
\author{XXXXX XXXXX$^{(a)}$, 
        YYYYY YYYYY$^{(b)}$, \\
        zzzzzz$^{(m)}$,
        yyyyyyy$^{(b)}$
        }

%xxxx$^{(1)}$, xxxxx$^{(1,3)}$}
\maketitle

\lyxaddress{{\small \vspace{-0.5cm}$^{(a)}$Escuela Polit\'{e}cnica Superior, Universidad Aut\'{o}noma de Madrid, 28049 Cantoblanco, Madrid, Spain.}}
\lyxaddress{{\small$^{(b)}$ The National Resource for Automated Molecular Microscopy, The Scripps Research Institute, La Jolla, CA 92037, USA}}
\lyxaddress{{\small$^{(c)}$ Baylor College of Medicine, Houston, Texas 77030, USA}}
\lyxaddress{{\small$^{(d)}$ Lawrence Berkeley National Laboratory, Berkeley, CA 94720, USA}}
\lyxaddress{{\small$^{(e)}$ Howard Hughes Medical Institute, Columbia University, NY 10032, USA}}
\lyxaddress{{\small$^{(f)}$ Purdue University, Biological Sciences, IN 47907-2054, USA}}
\lyxaddress{{\small$^{(g)}$ Biocomputing Unit, National Center for Biotechnology (CSIC), C/ Darwin, 3, Campus Universidad Aut\'{o}noma, 28049 Cantoblanco,
Madrid, Spain.}}
\lyxaddress{{\small$^{(h)}$ Biozentrum, University of Basel, CH - 4058 Basel, Switzerland}}
\lyxaddress{{\small$^{(i)}$ MRC-LMB, Cambridge CB2 0QH, UK. 01223 267000, United Kingdom}}
\lyxaddress{{\small$^{(j)}$IMPMC, Sorbonne Universit\'{e}s - CNRS UMR 7590, UPMC Univ Paris 6, MNHN, IRD UMR 206, 75005 Paris, France.}} 
\lyxaddress{{\small$^{(k)}$Roosevelt University, Department of Biological, Chemical, and Physical Sciences, 1400 N. Roosevelt Blvd., Schaumburg, IL 60173, USA.}}
\lyxaddress{{\small$^{(l)}$Laboratory of Structural Biology Research, National Institute of Arthritis, Musculoskeletal and Skin Diseases,
National Institutes of Health, Bethesda, MD 20892, USA}}
\lyxaddress{{\small$^{(m)}$Massachusetts Institute of Technology, USA}}
%\lyxaddress{{\small$^{(14)}$Massachusetts Institute of Technology, USA}}
% 
% \lyxaddress{{\small \vspace{-0.5cm}$^{(3)}$Bioengineering Lab., Universidad CEU San Pablo, Campus Urb.
% Montepríncipe s/n, 28668 Boadilla del Monte, Madrid, Spain.}}
%

%This document provides a preliminary analysis of the data collected by the CTF Challenge (http://i2pc.cnb.csic.es/3dembenchmark/LoadHome.htm). Figures/tables/plots/etc are located at the document's end. 
\setcounter{figure}{2}%I do not know why but figure counter starts with -1
\textbf{Corresponding Author:} %Roberto Marabini, Ph= 34 91 5854510, email: roberto@cnb.csic.es\\
%Escuela Polit\'{e}cnica Superior\\
%Universidad Aut\'{o}noma de Madrid\\
%28049 Cantoblanco\\
%Madrid\\
%Spain

\newpage
\input{abstract.tex}
\input{introduction.tex}
\input{overall.tex}
\input{workflow.tex}
\input{results.tex}
\input{conclusions.tex}
\input{acknowledgments.tex}
\appendix

\bibliographystyle{elsart-harv}
\bibliography{large}

\clearpage
%\input{listOfFigures.tex}

\newpage
%tables
%\input{numberImagesDataSet.tex}
%%%%%\input{markersLegend.tex}
%\input{numberImagesUpload.tex}
%\newpage

\end{document}