### 2018-04-02 6.acknowledgments.tex

\section{Acknowledgments}

The authors would like to acknowledge economical support from:

The Spanish Ministry of Economy and Competitiveness through Grants BIO2013-44647-R, BIO2016-76400-R(AEI/FEDER, UE) and AEI/FEDER BFU 2016 74868P, the Comunidad Aut\'{o}noma de Madrid through Grant: S2017/BMD-3817, European Union (EU) and Horizon 2020 through grant CORBEL (INFRADEV-1-2014-1, Proposal: 654248). The ``Knut \& Alice Wallenberg Foundation'', and ``A Pilot Facility development grant from Science for Life Laboratory''.

This work used the EGI Infrastructure and is co-funded by the EGI-Engage project (Horizon 2020) under Grant number 654142. European Union (EU) and Horizon 2020 through grant West-Life (EINFRA-2015-1, Proposal: 675858) European Union (EU) and Horizon 2020 through grant Elixir - EXCELERATE (INFRADEV-3-2015, Proposal: 676559) European Union (EU) and Horizon 2020 through grant iNEXT (INFRAIA-1-2014-2015, Proposal: 653706). The authors acknowledge the support and the use of resources of Instruct, a Landmark ESFRI project. This work has used the Cryo-EM Swedish National Facility funded by the Knut and Alice Wallenberg, Family Erling Persson and Kempe Foundations, SciLifeLab, Stockholm University and Ume\aa\ University.

We acknowledge Diamond for access and support of the Cryo-EM facilities at the UK national electron bio-imaging centre (eBIC), funded by the Wellcome Trust, MRC and BBSRC.

%\section{Project stability and funding}

%Scipion development is been done in the context of the Instruct Image Processing Cen-
%ter (I2PC, http://i2pc.cnb.csic.es). Instruct is the European initiative for Strategic
%Scientific Infrastructures (ESFRI) for Structural Biology. The I2PC started with the objective 
%to provide efficient support for research projects in Europe demanding expertise in 
%electron microscopy image processing. In order to support a large collection
%of projects, the first activities of the I2PC were oriented to design and implement the
%necessary software tools.

%At present the I2PC has this money and this people

%ScipionBox is part of the Scipion package which is provided 
%freely as open source software. Online documentation
%describing Scipion download and installation is available
%at \url{http://scipion.cnb.csic.es/m/download_form/} and \url{https://github.com/I2PC/scipion/wiki/How-to-Install}, respectively.

%It was recognized the need for a platform that could provide
%a friendly GUI on top of advanced image processing algorithms and, at the same time,
%assure standardization, traceability and reproducibility.
%\section{Things to mention if possible}
%\begin{itemize}
% \item Cloud
%\end{itemize}